# Time series decomposition

**Learning objectives:**

- how to split a time series into several components
- combine the trend and cycle into a single trend-cycle component
- extracting components from a time series


## Transformations and adjustments

To simplify the patterns in the historical data by removing known sources of variation some adjustments are required: 

- **calendar adjustments**
- **population adjustments** (data per person (or per thousand people, or per million people) rather than the total)
- **inflation adjustments** (such as price index adjustments)
- **mathematical transformations** (log/power/Box-Cox transformation, ...)

[Features by package](https://fabletools.tidyverts.org/reference/features_by_pkg.html)
[Guerrero's method for Box Cox lambda selection](https://feasts.tidyverts.org/reference/guerrero.html)
```{r message=FALSE, warning=FALSE}
library(fpp3)


lambda <- aus_production %>%
  features(Gas, features = guerrero) %>%
  pull(lambda_guerrero)


aus_production %>%
  autoplot(box_cox(Gas, lambda)) +
  labs(y = "",
       title = latex2exp::TeX(paste0(
         "Transformed gas production with $\\lambda$ = ",
         round(lambda,2))))+
  ggthemes::theme_pander()
```


## Time series components

**Time series** are made of three **components**: 

- a trend-cycle component $T_t$
- a seasonal component $S_t$
- a remainder component $R_t$


### Additive decomposition

When the variation around the **trend-cycle does not vary** with the level of the time series.

$$y_t=S_t+T_t+R_t$$

### Multiplicative decomposition


When the variation in the seasonal pattern, or the variation around the **trend-cycle**, appears to be **proportional** to the level of the time series.
 
 
$$y_t=S_t*T_t*R_t$$
An alternative can be a **log transformation**

$$log(y_t)=log(S_t)+log(T_t)+log(R_t)$$

### Example: Employment in the US retail sector


A nice introduction to the use of the **{fpp3} package** by the author is in the section *Example* at minute 13.11 here:
[Dr. Rob J. Hyndman - Ensemble Forecasts with {fable}](https://www.youtube.com/watch?v=nMy98aqQj8c)


**Data**
```{r}
us_employment%>%head
```


Decomposition with **STL decomposition method**: Seasonal and Trend decomposition using Loess.

[STL produces a smoother trend than classical decomposition methods due to the use of local polynomial regression.](https://arxiv.org/pdf/2204.10398.pdf)

Some interesting information about the **STL model**: 

- it is additive
- it is iterative and relies on the alternate estimation of the trend 
- the seasonal components are locally estimated scatterplot smoothing (Loess)
- it estimates nonlinear relationships
- the seasonal component is allowed to change over time
- it is composed of seasonal patterns estimated based on k consecutive seasonal cycles
- k controls how rapidly the seasonal component can change
- it is robust to outliers and missing data
- it is able to decompose time series with seasonality of any frequency, and provides implementation using numerical methods instead of mathematical modeling.

More info here: [R. B. Cleveland et al. (1990)](https://www.scb.se/contentassets/ca21efb41fee47d293bbee5bf7be7fb3/stl-a-seasonal-trend-decomposition-procedure-based-on-loess.pdf)

[from the book](https://otexts.com/fpp3/stl.html)

The command used is:

     model(stl = STL(<formula>))

```{r}
us_retail_employment <- us_employment %>%
  filter(year(Month) >= 1990, Title == "Retail Trade") %>%
  select(-Series_ID)


dcmp <- us_retail_employment %>%
  model(stl = STL(Employed))


components(dcmp) %>%
  autoplot()+
  ggthemes::theme_pander()
```

    ?components
    methods("components")


```{r}
components(dcmp) %>%
  as_tsibble()%>%
  select(Month,Employed,trend,season_adjust,remainder)%>%
  head
```


```{r message=FALSE, warning=FALSE}
components(dcmp) %>%# autoplot()
  as_tsibble() %>%
  autoplot(Employed, colour="grey80") +
  geom_line(aes(y=trend), colour = "navy",linewidth=2,alpha=0.5) +
  geom_line(aes(y=season_adjust), colour = "red",linewidth=0.3) +
  # geom_line(aes(y=remainder), colour = "blue") +
  labs(y = "Persons (thousands)",x="Year-Mon",
       title = "Total employment in US retail")+
  ggthemes::theme_pander()
```



## Classical decompositions

A classical decomposition method is the **moving average method** to estimate the trend-cycle. It originated in the 1920s and was widely used until the 1950s.


### Moving average smoothing

A moving average of order *m*: seasonal period  

$$m=2k+1$$

composed of seasonal patterns estimated based on **k** consecutive seasonal cycles, where $k$ controls how rapidly the component can change.

The estimate of the **trend-cycle** at time $t$ is obtained by averaging values of the time series within $k$ periods of $t$.

> the average eliminates some of the randomness

$$\hat{T_t}=\frac{1}{m}\sum_{j=-k}^k{y_{t+j}}$$
```{r}
aus_exports <- global_economy %>%
  filter(Country == "Australia")%>%
   mutate(`3-MA` = slider::slide_dbl(Exports, mean,
                                     .before = 1, # this is k
                                     .after = 1, 
                                     .complete = TRUE),
          `5-MA` = slider::slide_dbl(Exports, mean,
                                     .before = 2, # this is k
                                     .after = 2, 
                                     .complete = TRUE),
          `7-MA` = slider::slide_dbl(Exports, mean,
                                     .before = 3, # this is k
                                     .after = 3, 
                                     .complete = TRUE),
          `9-MA` = slider::slide_dbl(Exports, mean,
                                     .before = 4, # this is k
                                     .after = 4, 
                                     .complete = TRUE)
          ) %>%
  select(Year,Exports,`3-MA`,`5-MA`,`7-MA`,`9-MA`)%>%
  pivot_longer(cols = ends_with("MA"),names_to = "kma",values_to="values")

aus_exports %>% head
```

```{r warning=FALSE}
aus_exports %>%
  autoplot(Exports,color="darkred",size=0.2) +
  geom_line(aes(y = values,color=kma)) +
  labs(y = "% of GDP",x="Year",
       title = "Total Australian exports") +
  guides(colour = guide_legend(title = "series"))+
  ggthemes::scale_color_pander()+
  ggthemes::theme_pander()
```


### Moving averages of moving averages

When 2-MA follows a moving average of an even order (such as 4), it is called a “centered moving average of order 4”

The most common use of **centred moving averages** is for **estimating the trend-cycle from seasonal data**.


```{r}
beer_ma <- aus_production %>%
  filter(year(Quarter) >= 1992) %>%
  select(Quarter, Beer) %>%
  mutate(`4-MA` = slider::slide_dbl(Beer, mean,
                .before = 1, .after = 2, .complete = TRUE),
         `2x4-MA` = slider::slide_dbl(`4-MA`, mean,
                .before = 1, .after = 0, .complete = TRUE)
  )

beer_ma %>%head
```

### Weighted moving averages

Combinations of moving averages result in weighted moving averages.


Weights: $a_k=[\frac{1}{8},\frac{1}{4},\frac{1}{4},\frac{1}{4},\frac{1}{8}]$

k: $k=(m-1)/2$

$$\hat{T_t}=\sum_{j=-k}^k{a_jy_{t+j}}$$

### Additive decomposition

In **additive decomposition** is assumed that the seasonal component is constant from year to year. 


1. $$m=\left\{\begin{matrix}
 2\times m-MA &  \text{if m is even}\\ 
 m-MA &  \text{if m is odd}
\end{matrix}\right.$$

2.  De-trend: $y_t-\hat{T_t}$

3.  Adjust to ensure that they add to zero

4.  $R_t=y_t-\hat{T_t}-\hat{S_t}$

Formula:

    ?classical_decomposition
    model(classical_decomposition(<variable>, type = "additive"))

### Multiplicative decomposition

For **multiplicative seasonality**, the $m$ values that form the seasonal component are sometimes called the “seasonal indices”.

1. $$m=\left\{\begin{matrix}
 2\times m-MA &  \text{if m is even}\\ 
 m-MA &  \text{if m is odd}
\end{matrix}\right.$$

2.  De-trend: $y_t/\hat{T_t}$

3.  Adjust to ensure that they add to zero

4.  $R_t=y_t/(\hat{T_t}\hat{S_t})$


## Methods used by official statistics agencies

[US Census Bureau](https://www.bls.gov/), [Australian Bureau of Statistics](https://www.abs.gov.au/) and other **official statistics agencies** have developed their own decomposition procedures which are used for seasonal adjustment.

### X-11 method

It requires the [seasonal](https://bit.ly/seaspkg) package.
```{r}
x11_dcmp <- us_retail_employment %>%
  model(x11 = X_13ARIMA_SEATS(Employed ~ x11())) %>%
  components()


autoplot(x11_dcmp) +
  labs(title =
    "Decomposition of total US retail employment using X-11.")+
  ggthemes::theme_pander()
```

> The X-11 trend-cycle has captured the sudden fall in the data due to the 2007–2008 global financial crisis better than either of the other two methods.

```{r}
x11_dcmp %>%
  ggplot(aes(x = Month)) +
  geom_line(aes(y = Employed, colour = "Data")) +
  geom_line(aes(y = season_adjust,
                colour = "Seasonally Adjusted")) +
  geom_line(aes(y = trend, colour = "Trend")) +
  labs(y = "Persons (thousands)",
       title = "Total employment in US retail") +
  scale_colour_manual(
    values = c("gray", "#0072B2", "#D55E00"),
    breaks = c("Data", "Seasonally Adjusted", "Trend")
  )+
  ggthemes::theme_pander()
```

To help us visualise the variation in the seasonal component over time

```{r}
x11_dcmp %>%
  gg_subseries(seasonal)+
  ggthemes::theme_pander()
```


### The SEATS method

The **SEATS: Seasonal Extraction in ARIMA Time Series** developed at the Bank of Spain

```{r}
seats_dcmp <- us_retail_employment %>%
  model(seats = X_13ARIMA_SEATS(Employed ~ seats())) %>%
  components()


autoplot(seats_dcmp) +
  labs(title =
    "Decomposition of total US retail employment using SEATS")+
  ggthemes::theme_pander()
```



## Exercises

See this Rpub post: [https://rpubs.com/GehadGad/868705](https://rpubs.com/GehadGad/868705)




## Further reading

- [Author presentation](https://math.unm.edu/~lil/Stat581/6-decomposition.pdf)





## Meeting Videos

### Cohort 1

`r knitr::include_url("https://www.youtube.com/embed/IFv8OmMcfck")`

<details>
<summary> Meeting chat log </summary>

```
00:35:33	kevin kent:	Anomalize package which I mentioned. Has a method that uses STL and does anomaly detection on the remainder (by Matt Dancho) https://cran.r-project.org/web/packages/anomalize/vignettes/anomalize_quick_start_guide.html
```
</details>

`r knitr::include_url("https://www.youtube.com/embed/J2YWrDR3fF8")`

<details>
<summary> Meeting chat log </summary>

```
00:23:55	Ricardo Serrano:	https://rdrr.io/cran/fpp/man/melsyd.html
```
</details>


<!--
### Cohort 2

`r knitr::include_url("https://www.youtube.com/embed/URL")`

<details>
<summary> Meeting chat log </summary>

```
LOG
```
</details>
-->
