# The forecaster’s toolbox

**Learning objectives:**

- Learn how to apply NAIVE(y), SNAIVE(y) or RW(y ~ drift()) appropriately


## EXRCISE 1

Produce forecasts for the following series using whichever of NAIVE(y), SNAIVE(y) or RW(y ~ drift()) is more appropriate in each case:

- Australian Population (global_economy)
- Bricks (aus_production)
- NSW Lambs (aus_livestock)
- Household wealth (hh_budget)
- Australian takeaway food turnover (aus_retail)


```{r message=FALSE,warning=FALSE}
library(fpp3)
```


### Australian Population (global_economy)
```{r}

data(global_economy)

df <- global_economy %>%
  filter(Country=="Australia") %>%
  mutate(GDP_pop=GDP/Population)

df%>%head
```

```{r}
df %>%
  autoplot(Population)
```

```{r}

df %>% 
  select(Population,GDP_pop)%>%
  scale()%>%
  as_data_frame()%>%
  cbind(Year2=df$Year)%>%
  ggplot(aes(Year2,Population))+
  geom_point(shape=21,stroke=0.5)+
  geom_line()+
  geom_smooth(method = 'loess', se = FALSE, color = 'red',linewidth=0.3) + 
  geom_line(aes(Year2,GDP_pop))+
  scale_x_continuous(n.breaks = 10)+
  geom_text(aes(x=c(1990), y=c(0.5), label="Pop"),color="red")+
  geom_text(aes(x=c(2005), y=c(2.5), label="GDP/Pop"))+
  labs(title="Australian Population and GDP/Population")+
  theme_bw()
```

```{r}
df %>%
  model(
    `Naïve` = NAIVE(Population),
    `Seasonal naïve` = SNAIVE(Population),
    Drift = RW(log(Population) ~ drift())
  ) %>%
    forecast(h = c(10,20)) %>%
    autoplot(df |> filter(!is.na(Population)),
    point_forecast = lst(mean, median)
  )
```
```{r}
fit_pop_naive <- df %>%
  model( `Naïve` = NAIVE(Population))
```

```{r}
fit_pop_naive |> gg_tsresiduals()
```
```{r}
fit_pop_drift <- df %>%
  model(Drift = RW(log(Population) ~ drift()))
  

fit_pop_drift |> gg_tsresiduals()
```


### Bricks (aus_production)
```{r}
data(aus_production)
aus_production%>%head
```


```{r}
aus_production%>%
  autoplot(Bricks)
```


```{r}
bricks_fit <- aus_production |>
  model(
    Mean = MEAN(Bricks),
    `Naïve` = NAIVE(Bricks),
    `Seasonal naïve` = SNAIVE(Bricks),
    Drift = RW(Bricks ~ drift())
  )

bricks_fc <- bricks_fit |>
  forecast(h = 10)

bricks_fc |>
  autoplot(
    aus_production,
    level = NULL
  ) +
  labs(
    y = "",
    title = "Forecasts for quarterly Bricks production"
  ) +
  guides(colour = guide_legend(title = "Forecast"))
```

### NSW Lambs (aus_livestock)
```{r}
data(aus_livestock)%>%head
```


### Household wealth (hh_budget)
```{r}
data(hh_budget)%>%head
```


### Australian takeaway food turnover (aus_retail)
```{r}
data(aus_retail)%>%head
```





## Meeting Videos

### Cohort 1

`r knitr::include_url("https://www.youtube.com/embed/z1DY5xzHUKY")`

<details>
<summary> Meeting chat log </summary>

```
00:43:42	Ricardo Serrano:	In modeltime package, there is a pull_modeltime_residuals() function that looks promising for residual diagnostics.
00:48:07	Kevin Kent:	cool!
01:10:15	Mikhael Manurung:	Noted!
01:13:51	Federica Gazzelloni:	thank you
```
</details>


### Cohort 2

`r knitr::include_url("https://www.youtube.com/embed/8uwXytWn5HU")`

`r knitr::include_url("https://www.youtube.com/embed/gDqmMd9x2FA")`

`r knitr::include_url("https://www.youtube.com/embed/-RghEM6seGU")`

<details>
<summary> Meeting chat log </summary>

```
00:12:41	Ricardo Serrano: Random walk definition https://www.quantstart.com/articles/White-Noise-and-Random-Walks-in-Time-Series-Analysis/#:~:text=A%20random%20walk%20is%20another,random%20step%20up%20or%20down.
00:12:58	Ricardo Serrano: IS your time series a random walk https://machinelearningmastery.com/gentle-introduction-random-walk-times-series-forecasting-python/#:~:text=A%20random%20walk%20is%20one,in%20stock%20prices%20are%20unpredictable.
```
</details>
